\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-02 draft.tex
move assumptions above sec:twt

One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain. Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior. While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.

@@

Therefore, in line with both the theoretical and empirical literature, I make the following assumption:

@@

Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$. I make the following further assumptions on the weight on import-industry profits:

Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.

Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.

>>>>

One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain. Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior. I make the following two basic assumptions on the weight on import-industry profits:

Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace. Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.

While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.

@@

Therefore, in line with both the theoretical and empirical literature, I make the following, final assumption on $\ga(e,\ta)$:

@@

Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$. Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}. More restrictions could be added to deliver further results.
